\large{Vipassana for Hackers} \\ \Huge{Paper Two: The Brain} \\ \large\textit{Version 0.2}

### 2019-10-09 paper-two/paper-two.tex
adding homunculus diagram to pdf
update todo what does this mean?

\todo{Are the labels on either the SVG or the MRI image reversed?}

@@

Over the final seven days of the course, as awareness gets smoother and observations get more objective, this
functional connectivity between the ACC and the OFC slowly starts changing. While
continuing to pay attention to constantly changing bodily sensations the goal remains: avoid
assigning evaluations (and therefore, emotions) to them. Without assigning an
emotional response, one learns to observe with a
non-judgmental attitude, particularly toward unpleasant stimuli. The OFC still fires
but the Limbic System is quiesced and, in turn, reactions become smaller and less
frequent. This snowball effect almost mirrors the spiral of brain behaviours
exhibited early in the course. Early in the course, the CEN's normal mode of
operation is in play, which is to experience unpleasant stimuli and react (desiring to rid the body of
them), which in turn causes the unpleasant stimuli to appear all the more
unpleasant. Once a measure of objectivity is reached, however, the brain's function
snowballs in the other direction. It is as if both conscious and unconscious awarenesses are slowly aligning
until there are occasions of experience when it feels as if there is nothing but detached awareness moving
through the body (or some parts of the body). This is not at all akin to the
detachment of sleep or intoxication, however, and the senses are not dulled. In
addition to a feeling of emotional detachment, the AIC and the Somatosensory
Cortex, which are involved in sensory awareness, were \textbf{more} active, not
less. As the ACC, OFC, and Limbic System are calmed and attention is narrowed
further, the sensations become progressively more intense.

Perhaps the strangest experience during a Vipassana course is that of shifts in the
sense of self. The ``self'', as a subjective agent in space and time, changes. At
first, this change is only very slight. As a student closes her eyes and begins to
meditate, it may feel as though
the body has shifted in space or rotated in different directions. During these times,
the vmPFC and PCC become less active. \todo{So? What does this mean?} As the meditations get deeper the DMN
occasionally quiets completely (perhaps only for a short duration) while the AIC
fires more rapidly, leading to alternative networks of consciousness
arising. \todo{What does this mean? What is an example of ``alternative networks of consciousness''?} It is
as if losing the influence of one's own narrative leads to new insights that are
usually kept from consciousness. It has been found that experiencing this, even for a
few seconds, can lead to lasting changes long after the activity of meditation itself
has stopped. \cite{alteredtraits} In the moment, these experiences can be quite
similar to being on certain hallucinogenic drugs or even, in some cases, to ecstatic
seizures where the abnormal activity of the Anterior Insular Cortex leads to
heightened self-awareness, feelings of bliss, and a lack of ambiguity. \cite{cortexbliss}

>>>>

@@

Over the final seven days of the course, as awareness gets smoother and observations get more objective, this
functional connectivity between the ACC and the OFC slowly starts changing. While
continuing to pay attention to constantly changing bodily sensations the goal remains: avoid
assigning evaluations (and therefore, emotions) to them. Without assigning an
emotional response, one learns to observe with a
non-judgmental attitude, particularly toward unpleasant stimuli. The ACC is still up-regulated 
but the OFC and the Limbic System may acquiesce and, in turn, reactions become smaller and less
frequent. This snowball effect almost mirrors the spiral of brain behaviours
exhibited early in the course. Early in the course, the CEN's normal mode of
operation is in play, which is to experience unpleasant stimuli and react (desiring to rid the body of
them), which in turn causes the unpleasant stimuli to appear all the more
unpleasant. Once a measure of objectivity is reached, however, the brain's function
snowballs in the other direction. It is as if both conscious and unconscious awarenesses are slowly aligning
until there are occasions of experience when it feels as if there is nothing but detached awareness moving
through the body (or some parts of the body). This is not at all akin to the
detachment of sleep or intoxication, however, and the senses are not dulled. In
addition to a feeling of emotional detachment, the ACC, AIC and the Somatosensory
Cortex, which are involved in sensory awareness, were \textbf{more} active, not
less. As the OFC, and Limbic System are calmed and attention is narrowed
further, the sensations become progressively more intense.

Perhaps the strangest experience during a Vipassana course is that of shifts in the
sense of self. The ``self'', as a subjective agent in space and time, changes. At
first, this change is only very slight. As a student closes her eyes and begins to
meditate, it may feel as though the body has shifted in space or rotated in different direction, 
due to the temporary down regulation of the prefrontal cortex (don't worry about it, keep observing
sensations). \todo{So? What does this mean?} As the meditations get deeper the DMN
occasionally quiets completely (perhaps only for a short duration) while the AIC
fires more rapidly, leading to alternative networks of consciousness
arising. \todo{What does this mean? What is an example of ``alternative networks of consciousness''?} It is
as if losing the influence of one's own narrative leads to new insights that are
usually kept from consciousness. It has been found that experiencing this, even for a
few seconds, can lead to lasting changes long after the activity of meditation itself
has stopped. \cite{alteredtraits} In the moment, these experiences can be quite
similar to being on certain hallucinogenic drugs or even, in some cases, to ecstatic
seizures where the abnormal activity of the Anterior Insular Cortex leads to
heightened self-awareness, feelings of bliss, and a lack of ambiguity. \cite{cortexbliss}
